Enabling GPU Virtualization in Cloud Environments

### 2015-10-23 paper.tex

\begin{figure*}[ht]
\centering
\subfigure[5 remote GPUs over 1GbE.]{%
\includegraphics[width=0.31\linewidth]{images/aws3.png}
\label{subfig:aws1}}
\quad
\subfigure[8 remote GPUs over 10GbE.]{%
\includegraphics[width=0.31\linewidth]{images/aws2.png}
\label{subfig:aws2}}
\subfigure[13 remote GPUs over 1GbE.]{%
\includegraphics[width=0.31\linewidth]{images/aws1.png}
\label{subfig:aws3}}
\caption{Different configurations where an AWS instance acting as a client is using several remote GPUs.}
\label{fig:aws}
\end{figure*}

@@

\subsection{Experimental Results}
The following tests were executed on 2 sets of nodes interconnected by a 1Gb Ethernet network.

The first set, in charge of providing the cloud environment,
consisted of 3 nodes, each equipped with an Intel Xeon E7420 quadcore processor at
2.13 GHz, and 16 Gbytes of DDR2 RAM memory at 667 MHz.
To deploy an infrastructure as a cloud service, we used OpenStack Icehouse version and QEMU/KVM 0.12.1 as the hypervisors.
A fully-featured OpenStack deployment requires at least three nodes: one controller node that manages the compute nodes 
where the VMs are hosted, one network node that manages the logic
virtual network for the VMs, and one or more compute nodes that 
run the hypervisor and virtual machines.

The second set, composed by 4 nodes, were auxiliary servers with a Tesla C1060 GPU each one. 
Their system operates a Centos 6.6; and the GPUs use CUDA 6.5. 
The GPU virtualization support is based on rCUDA v5.0.

We have designed 6 different set-ups which can be divided into 2 groups: “exclusive GPUs” and “shared GPUs”. 
The “exclusive” mode provides, at most, 4 accelerators. 
Whilst the number of available GPU in “shared” mode will depend on the partition size. 
In this particular case, we have halved the GPU memory, having as a result 8 partitions that can be addressed as independent GPUs.
For each group, we have deployed virtual clusters of 1, 2 and 3 nodes where the application processes will be executed. 
The instances are lanched with the Openstack predifined flavor “m1.medium”, which determines a configuration of VMs consisting of 2 cores and 4096 MB of memory.

@@

\section{\uppercase{Conclusions}}
\label{sec:conclusions}
We have seen a thoroughly study of the possibilities offered by AWS in everything to do with GPUs. Its limitations drove us to deploy our own private cloud to have more flexibility when working with these accelerators.
So that, we have presented an upgrade for Openstack which easily permits to create GPU-instances and manage the physical GPUs to make more profit of them.

Leaving aside the performance of the executions which we knew in advance that would be poor due to the interconnection network features, we have created new operation modes. They have brought interesting new ways of using GPUs in situations where the simple fact of having accessing a GPU is more important than having a powerful GPU to boost the performance.

\section{\uppercase{Future work}}
\label{sec:future}
The first task that we should have keep in mind after this study is to upgrade the interconnection network to one more focused on HPC.
So that, porting the setup and the tests to an infrastructure with an Infiniband network would shed light on the viability of this kind of solutions for HPC.
The same thought, would guide us to try other Cloud vendors which more support for HPC.

Looking for real situations where the performance is not so important as the flexibility to provide more accelerators, 
would drive us to take the first steps through tools for easily deploying programming computer labs with the necessity for GPUs.

Other interesting approach would be to make the system energy-aware. 
Provided that a GPU could throw power information, a similar study based on the energy consumption could be done.

Finally, to bring in more strategies to decide where a remote GPUs is created and assigned to a physical device, would be an interesting task to carry out. 
With more scheduling policies the flexibility given by the GPGPU module for OpenStack would increase.

>>>>

\begin{figure*}[ht]
\centering
\subfigure[5 remote GPUs over 1GbE.]{%
\includegraphics[width=0.315\linewidth]{images/aws3.pdf}
\label{subfig:aws1}}
\quad
\subfigure[8 remote GPUs over 10GbE.]{%
\includegraphics[width=0.315\linewidth]{images/aws2.pdf}
\label{subfig:aws2}}
\subfigure[13 remote GPUs over 1GbE.]{%
\includegraphics[width=0.315\linewidth]{images/aws1.pdf}
\label{subfig:aws3}}
\caption{Different configurations where an AWS instance acting as a client is using several remote GPUs.}
\label{fig:aws}
\end{figure*}

@@

\subsection{Experimental Results}
The following tests were executed on 2 sets of nodes interconnected by a 1Gb Ethernet network.
All the nodes were equipped with an Intel Xeon E7420 quadcore processor at 2.13 GHz, and 16 Gbytes of DDR2 RAM memory at 667 MHz.

The first set, in charge of providing the cloud environment,
consisted of 3 nodes.
To deploy an infrastructure as a cloud service, we used OpenStack Icehouse version and QEMU/KVM 0.12.1 as the hypervisor.
A fully-featured OpenStack deployment requires at least three nodes: a controller node manages the compute nodes 
where the VMs are hosted, a network node that manages the logic virtual network for the VMs, and one or more compute nodes that 
run the hypervisor and VMs.

The second set, composed of 4 nodes, were auxiliary servers with a Tesla C1060 GPU each one. 
The OS was a Centos 6.6; and the GPUs used CUDA 6.5. 
The GPU virtualization support was based on rCUDA v5.0.

We have designed 6 different set-ups which can be divided into 2 groups: “exclusive GPUs” and “shared GPUs”. 
The “exclusive” mode provides, at most, 4 accelerators. 
Whilst the number of available GPU in “shared” mode will depend on the partition size. 
In this particular case, we have halved the GPU memory, having as a result 8 partitions that can be addressed as independent GPUs.
For each group, we have deployed virtual clusters of 1, 2 and 3 nodes where the application processes will be executed. 
The instances are lanched with the OpenStack predifined flavor “m1.medium”, which determines a configuration of VMs consisting of 2 cores and 4096 MB of memory.

@@

\section{\uppercase{Conclusions}}
\label{sec:conclusions}
We have presented a complete study of the possibilities offered by AWS when it comes to GPUs. 
The constraints imposed by this service motivated us to deploy our own private cloud, in order to gain flexibility when working with these accelerators.
For this purpose, we have introduced an extension of OpenStack which easily permits to create GPU-instances and manage the physical GPUs to better profit from them.

As we expected, due to the limited bandwidth of the interconnects used in the experimentation, the performance observed for the GPU virtualized scenarios in the test was quite low.
On the other hand we have created new operation modes that open interesting new ways of using GPUs in situations where the having access to a GPU is more important than having a powerful GPU to boost the performance.

\section{\uppercase{Future work}}
\label{sec:future}
The first in the list of pending work is the upgrade of the interconnection network to one more prone on HPC.
Porting the setup and the tests to an infrastructure with an Infiniband network will shed light on the viability of this kind of solutions for HPC.
The same thought, motivates us to try other Cloud vendors which better support for HPC.

Looking for real situations where performance is less important than flexibility will drive us 
to tools for easily deploying GPU-programming computer labs.

Finally, an interesting future mode is to design new strategies in order to decide where a remote GPUs is created and assigned to a physical device
In particular, innovate scheduling policies can enhance the flexibility given by the GPGPU module for OpenStack.
